---
title: "Exercise Lesson 7"
author: "Janna & Stefan"
date: "Wednesday, January 14, 2015"
output: html_document
---

This is our result of the exercise of Lesson 7

First we load the libraries:
```{r}
library(raster)
```

Next we load the data from the data folder:
```{r}
load("data/GewataB1.rda")
load("data/GewataB2.rda")
load("data/GewataB3.rda")
load("data/GewataB4.rda")
load("data/GewataB5.rda")
load("data/GewataB7.rda")
load("data/vcfGewata.rda")
```

Then we remove the unrealistic values from the vcf file (which contains tree cover):
```{r}
vcfGewata[vcfGewata > 100] <- NA
```

After that we put the bands together in a brick...
```{r}
gewata <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7, vcfGewata)
gewata_novcf <- brick(GewataB1, GewataB2, GewataB3, GewataB4, GewataB5, GewataB7)
```

... and we plot the bands in pairs to show correlations:
```{r}
pairs(gewata)
```

Now we choose the bands with the highest correlation and we plot them seperately:
```{r}
pairs(brick(GewataB7, vcfGewata))
pairs(brick(GewataB3, vcfGewata))
```

Now we create the linear model, which predicts the VCF with the reflections of the different bands.
In the summary a column named 'Pr(>|t|)' is given. Small values mean that the probability that this correlation results from coincidence is really small, so the bands with the small p-values are important: these are all the bands except for band 7.
```{r}
valuetable <- as.data.frame(getValues(gewata))
modelLM <- lm(formula = vcf2000Gewata ~ gewataB1 + gewataB2 + gewataB3 + gewataB4 + gewataB5 + gewataB7, data = valuetable)
summary(modelLM)
```

Now we make the predicted tree cover raster and plot it:
```{r fig.width=10}
predLC <- predict(gewata_novcf, model=modelLM, na.rm=TRUE)
predLC[predLC < 0] <- NA

plot(predLC, main="Predicted forest coverage", zlim=c(0,100))
plot(gewata$vcf2000Gewata, main="Actual forest coverage", zlim=c(0,100))
```

Then we calculate the RMSE by using rasterlayers:
```{r}
sqdif <- (gewata$vcf2000Gewata-predLC)^2
sqdifdf <- as.data.frame(sqdif)
RMSE_rasterlayer<-sqrt(mean(sqdifdf$layer, na.rm=TRUE))
RMSE_rasterlayer
```

It is also possible to calculate the RMSE by doing the following:
```{r}
RMSE_cellstats <- sqrt(cellStats(((gewata$vcf2000Gewata-predLC)^2), stat='mean'))
RMSE_cellstats
```

The last part is to calculate the difference between the actual and predicted tree coverage per class
```{r}
differenceTC<-(gewata$vcf2000Gewata-predLC)^2

load("data/trainingPoly.rda")
trainingPoly@data$Code <- as.numeric(trainingPoly@data$Class)
ndvi <- overlay(GewataB4, GewataB3, fun=function(x,y){(x-y)/(x+y)})
classes <- rasterize(trainingPoly, ndvi, field='Code')

averagediff<-zonal(differenceTC, classes)
RMSEClasses<-sqrt(averagediff[,2])
RMSEClasses
```

This results in an RMSE=8.87 for cropland, an RMSE=4.99 for forest and an RMSE=11.91 for wetland. This means that the difference between the predicted and actual tree cover is the biggest for wetland.
